Identification of high-yielding soybean lines with exceptional seed composition qualities

### 2022-05-31 doc/manuscript/manuscript.Rmd
Added to the manuscript document. Also made some changes to the summary tables and added a step to make summary tables for smaller selections of genotypes.

```{r setup, warning=FALSE,echo = FALSE, message=FALSE}
library(here)
source(here("packages.R"))
tar_load(c(example_table, 
           yield_summary_tables))
```

@@

To meet this goal, two recombinant inbred line (RIL) oil mapping populations were screened for lines which showed promising combinations of yield and seed composition traits. Successive rounds of selection were conducted to identify and characterize lines with high values for yield as well as protein and oil composition were performed between 2018 and 2021 to identify soybean lines with elevated yield as well as the valuable seed composition traits when compared with existing check cultivars.

@@

Models were fit using the gamem_met function from the metan package[@olivotoMetanPackageMultienvironment2020]. Least-square means (LS Means) for each genotype and trait were calculated using the above model using the emmeans package[@lenthEmmeansEstimatedMarginal2022] in R. The emmeans package was also used to calculate contrasts as a post-hoc test to compare RIL means to check means. Pearson correlation coefficients between each phenotype were calculated and visualized with the metan package.

@@

Several RILs in each of the tests had yield that was comparable to that of their check cultivars. This was determined both quanititatively through the use of contrasts between the mean of each RIL with the average of the yield check means, and qualitatively by inspecting the standard errors on the estimation of the genotype marginal means for yield. Among these genotypes with yield comparable to that of the check cultivars were several genotypes with

```{r test1_summary_table, results = "asis", echo = FALSE}

print(yield_summary_tables$`Jay Test 1`)

>>>>

```{r setup, warning=FALSE,echo = FALSE, message=FALSE}
library(here)
source(here("packages.R"))
tar_load(c(example_table, 
           yield_summary_tables, 
           elite_genotype_tables))
```

@@

To meet this goal, two recombinant inbred line (RIL) oil mapping populations were screened for lines which showed promising combinations of yield and seed composition traits. Successive rounds of selection were conducted to identify and characterize lines with high values for yield as well as protein and oil composition were performed between 2018 and 2021 to identify soybean lines with yield comparable to existing check cultivars and protein and or oil composition superior to that of the check cultivars.

@@

Models were fit using the gamem_met function from the metan package[@olivotoMetanPackageMultienvironment2020]. Least-square means (LS Means) for each genotype and trait were calculated using the above model using the emmeans package[@lenthEmmeansEstimatedMarginal2022] in R. The emmeans package was also used to calculate contrasts as a post-hoc test to compare RIL phenotype means to check means. Pearson correlation coefficients between each phenotype were calculated and visualized with the metan package.

@@

Several RILs in each of the tests had yield that was comparable to that of their check cultivars. This was determined both quanititatively through the use of contrasts between the mean of each RIL with the average of the yield check means, and qualitatively by inspecting the standard errors on the estimation of the genotype marginal means for yield. No genotypes had yield that was quantitatively higher than that of the yield checks, however many had comparable yield. Many of these genotypes with comparable yield also had protein content that was superior to the check average in each test. No genotypes had oil that was superior to the check cultivars, but some had comparable yield and oil content as well as superior protein content. These genotypes with comparable yield and superior protein also had desirable agronomic qualities such as low lodging and good seed quality which would make them ideal candidates for future breeding material.

## Genotypes with omparable yield and oil and superior protein
Four genotypes had yield and oil that was similar to that of the yield checks, and seed protein that was greater than the average of the checks. These genotypes are N18-1635 from Test 1 and genotypes N18-1627, N18-1643, and N18-1783 from Test 2.

```{r elite_geno_table, results = "asis", echo = FALSE}

print(elite_genotype_tables$elite_genos)

```{r high_po_table, results = "asis", echo = FALSE}

print(elite_genotype_tables$po_top_ten)

### 2022-05-31 doc/manuscript/manuscript.tex
Added to the manuscript document. Also made some changes to the summary tables and added a step to make summary tables for smaller selections of genotypes.

Models were fit using the gamem\_met function from the metan
package\citep{olivotoMetanPackageMultienvironment2020}. Least-square
means (LS Means) for each genotype and trait were calculated using the
above model using the emmeans
package\citep{lenthEmmeansEstimatedMarginal2022} in R. The emmeans
package was also used to calculate contrasts as a post-hoc test to
compare RIL means to check means. Pearson correlation coefficients
between each phenotype were calculated and visualized with the metan
package.

@@

Several RILs in each of the tests had yield that was comparable to that
of their check cultivars. This was determined both quanititatively
through the use of contrasts between the mean of each RIL with the
average of the yield check means, and qualitatively by inspecting the
standard errors on the estimation of the genotype marginal means for
yield. Among these genotypes with yield comparable to that of the check
cultivars were several genotypes with

\caption{Marginal means and rankings of genotypes in Jay Test 1 for seed yield, seed oil, and seed protein.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{ccccccccccccc}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Protein} & \multicolumn{4}{c}{Yield} & \multicolumn{4}{c}{Oil} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9} \cmidrule(l{3pt}r{3pt}){10-13}
Genotype & Value\textbackslash{}textsuperscript\{*\} & Rank\textbackslash{}\textbackslash{}textsuperscript\{\textbackslash{}\textbackslash{}dag\} & Test Average\textbackslash{}\textbackslash{}textsuperscript\{\textbackslash{}\textbackslash{}ddag\} & Check Average\textbackslash{}\textbackslash{}textsuperscript\{\textbackslash{}\textbackslash{}S\} & Value & Rank & Test Average & Check Average & Value & Rank & Test Average & Check Average\\
\midrule
LMN09-119 & 51.37 (125.5\%) & 1 &  &  & 39.36 (87.4\%) & 18 &  &  & 18.01 (80.2\%) & 20 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1763 & 47.82 (116.8\%) & 2 &  &  & 39.45 (87.6\%) & 17 &  &  & 19.58 (87.2\%) & 18 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1855 & 47.52 (116.1\%) & 3 &  &  & 43.17 (95.8\%) & 10 &  &  & 20.6 (91.8\%) & 14 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1595 & 47.36 (115.7\%) & 4 &  &  & 43.52 (96.6\%) & 9 &  &  & 19.52 (86.9\%) & 19 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
Dunphy & 41.71 (101.9\%) & 18 &  &  & 47.33 (105.1\%) & 1 &  &  & 21.52 (95.9\%) & 4 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
Dilday & 41.22 (100.7\%) & 19 &  &  & 46.78 (103.8\%) & 2 &  &  & 22.33 (99.5\%) & 2 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1632-2 & 46.67 (114\%) & 5 &  &  & 46.34 (102.9\%) & 5 &  &  & 20.8 (92.7\%) & 11 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1632-1 & 46.64 (113.9\%) & 6 &  &  & 45.73 (101.5\%) & 6 &  &  & 20.76 (92.5\%) & 12 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1620 & 43.63 (106.6\%) & 15 &  &  & 46.54 (103.3\%) & 3 &  &  & 21.4 (95.3\%) & 6 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1635 & 43.61 (106.5\%) & 16 &  &  & 46.45 (103.1\%) & 4 &  &  & 21.72 (96.7\%) & 3 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N09-09 & 46.31 (113.1\%) & 7 &  &  & 41.61 (92.4\%) & 14 &  &  & 19.77 (88.1\%) & 17 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1731 & 46.1 (112.6\%) & 8 &  &  & 43.16 (95.8\%) & 11 &  &  & 20.49 (91.3\%) & 15 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1674 & 45.19 (110.4\%) & 9 &  &  & 44.26 (98.2\%) & 7 &  &  & 21.21 (94.5\%) & 9 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1682 & 45.11 (110.2\%) & 10 &  &  & 43.64 (96.9\%) & 8 &  &  & 20.76 (92.5\%) & 13 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1572 & 45.08 (110.1\%) & 11 &  &  & 38.58 (85.6\%) & 19 &  &  & 21.19 (94.4\%) & 10 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1751 & 44.8 (109.4\%) & 12 &  &  & 41.59 (92.3\%) & 15 &  &  & 20.44 (91\%) & 16 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1956 & 44.55 (108.8\%) & 13 &  &  & 34.79 (77.2\%) & 20 &  &  & 21.51 (95.8\%) & 5 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1604 & 43.83 (107.1\%) & 14 &  &  & 43.15 (95.8\%) & 12 &  &  & 21.34 (95.1\%) & 8 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
N18-1641 & 43.33 (105.8\%) & 17 &  &  & 41.99 (93.2\%) & 13 &  &  & 21.36 (95.1\%) & 7 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
\cmidrule{10-11}
NC-Raleigh & 39.91 (97.5\%) & 20 & \multirow{-20}{*}{\centering\arraybackslash 45.09} & \multirow{-20}{*}{\centering\arraybackslash 40.94} & 41.04 (91.1\%) & 16 & \multirow{-20}{*}{\centering\arraybackslash 42.92} & \multirow{-20}{*}{\centering\arraybackslash 45.05} & 23.5 (104.7\%) & 1 & \multirow{-20}{*}{\centering\arraybackslash 20.89} & \multirow{-20}{*}{\centering\arraybackslash 22.45}\\
\bottomrule
\multicolumn{13}{l}{\rule{0pt}{1em}\textsuperscript{*} Observed marginal mean of the phenotype (value of the phenotype/check average * 100\%).}\\
\multicolumn{13}{l}{\rule{0pt}{1em}\textsuperscript{\dag} Rank of the genotype within the particular phenotype.}\\
\multicolumn{13}{l}{\rule{0pt}{1em}\textsuperscript{\ddag} The average value of the phenotype for all lines in the test.}\\
\multicolumn{13}{l}{\rule{0pt}{1em}\textsuperscript{\S} The average value of the phenotype for the checks in the test.}\\
\end{tabular}}
\end{table}

@@

\begin{figure}[H]
\centering
\includegraphics[width=10 cm]{C:/Users/Jay/Desktop/Documents/R/Jay_yield/exports/plots/example_plot.pdf}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

>>>>

Models were fit using the gamem\_met function from the metan
package\citep{olivotoMetanPackageMultienvironment2020}. Least-square
means (LS Means) for each genotype and trait were calculated using the
above model using the emmeans
package\citep{lenthEmmeansEstimatedMarginal2022} in R. The emmeans
package was also used to calculate contrasts as a post-hoc test to
compare RIL phenotype means to check means. Pearson correlation
coefficients between each phenotype were calculated and visualized with
the metan package.

@@

Several RILs in each of the tests had yield that was comparable to that
of their check cultivars. This was determined both quanititatively
through the use of contrasts between the mean of each RIL with the
average of the yield check means, and qualitatively by inspecting the
standard errors on the estimation of the genotype marginal means for
yield. No genotypes had yield that was quantitatively higher than that
of the yield checks, however many had comparable yield. Many of these
genotypes with comparable yield also had protein content that was
superior to the check average in each test. No genotypes had oil that
was superior to the check cultivars, but some had comparable yield and
oil content as well as superior protein content. These genotypes with
comparable yield and superior protein also had desirable agronomic
qualities such as low lodging and good seed quality which would make
them ideal candidates for future breeding material.

\hypertarget{genotypes-with-omparable-yield-and-oil-and-superior-protein}{%
\subsection{Genotypes with omparable yield and oil and superior
protein}\label{genotypes-with-omparable-yield-and-oil-and-superior-protein}}

Four genotypes had yield and oil that was similar to that of the yield
checks, and seed protein that was greater than the average of the
checks. These genotypes are N18-1635 from Test 1 and genotypes N18-1627,
N18-1643, and N18-1783 from Test 2.

\begin{table}[H]

\caption{Soybean genotypes with yield and seed oil comparable to check cultivars, and seed protein superior to check cultivars.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llrlllrll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Protein + Oil} & \multicolumn{4}{c}{Yield} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
Genotype & Value & Rank & Test Average & Check Average & Value & Rank & Test Average & Check Average\\
\midrule
N18-1635 & 65.32 (103\%) & 13 & 65.98 & 63.39 & 46.45 (103.1\%) & 4 & 42.92 & 45.05\\
\cmidrule{1-9}
N18-1783 & 67.38 (105.1\%) & 8 &  &  & 37.63 (86.6\%) & 19 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
N18-1627 & 66.52 (103.7\%) & 10 &  &  & 46.01 (105.9\%) & 8 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
N18-1643 & 65.8 (102.6\%) & 13 & \multirow{-3}{*}{\raggedright\arraybackslash 66.4} & \multirow{-3}{*}{\raggedright\arraybackslash 64.14} & 40.54 (93.3\%) & 16 & \multirow{-3}{*}{\raggedright\arraybackslash 43.65} & \multirow{-3}{*}{\raggedright\arraybackslash 43.44}\\
\bottomrule
\multicolumn{9}{l}{\rule{0pt}{1em}\textsuperscript{*} The genotype name.}\\
\multicolumn{9}{l}{\rule{0pt}{1em}\textsuperscript{\dag} The genotype marginal mean for the phenotype (value divided by check average).}\\
\multicolumn{9}{l}{\rule{0pt}{1em}\textsuperscript{\ddag} The ranking of this genotype for the phenotype within its test.}\\
\multicolumn{9}{l}{\rule{0pt}{1em}\textsuperscript{\S} The average phenotype value for all genotypes in the test.}\\
\multicolumn{9}{l}{\rule{0pt}{1em}\textsuperscript{\P} The average value of the checks in the test.}\\
\end{tabular}}
\end{table}

\begin{table}[H]

\caption{Top 10 soybean genotypes on the basis of protein plus oil content.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llrlllrll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Protein + Oil} & \multicolumn{4}{c}{Yield} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
Genotype & Value & Rank & Test Average & Check Average & Value & Rank & Test Average & Check Average\\
\midrule
N18-1855 & 68.13 (107.5\%) & 2 &  &  & 43.17 (95.8\%) & 10 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
N18-1632-2 & 67.47 (106.4\%) & 3 &  &  & 46.34 (102.9\%) & 5 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
N18-1763 & 67.4 (106.3\%) & 5 &  &  & 39.45 (87.6\%) & 17 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
N18-1595 & 66.87 (105.5\%) & 6 &  &  & 43.52 (96.6\%) & 9 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
N18-1731 & 66.59 (105\%) & 7 &  &  & 43.16 (95.8\%) & 11 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
N18-1674 & 66.4 (104.7\%) & 8 & \multirow{-6}{*}{\raggedright\arraybackslash 65.98} & \multirow{-6}{*}{\raggedright\arraybackslash 63.39} & 44.26 (98.2\%) & 7 & \multirow{-6}{*}{\raggedright\arraybackslash 42.92} & \multirow{-6}{*}{\raggedright\arraybackslash 45.05}\\
\cmidrule{1-9}
N18-1575 & 69.04 (107.6\%) & 1 &  &  & 42.49 (97.8\%) & 12 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
N18-1761 & 68.48 (106.8\%) & 3 &  &  & 45.74 (105.3\%) & 9 &  & \\
\cmidrule{1-3}
\cmidrule{6-7}
N18-1769 & 67.95 (105.9\%) & 6 & \multirow{-3}{*}{\raggedright\arraybackslash 66.4} & \multirow{-3}{*}{\raggedright\arraybackslash 64.14} & 37.88 (87.2\%) & 18 & \multirow{-3}{*}{\raggedright\arraybackslash 43.65} & \multirow{-3}{*}{\raggedright\arraybackslash 43.44}\\
\bottomrule
\multicolumn{9}{l}{\rule{0pt}{1em}\textsuperscript{*} The genotype name.}\\
\multicolumn{9}{l}{\rule{0pt}{1em}\textsuperscript{\dag} The genotype marginal mean for the phenotype (value divided by check average).}\\
\multicolumn{9}{l}{\rule{0pt}{1em}\textsuperscript{\ddag} The ranking of this genotype for the phenotype within its test.}\\
\multicolumn{9}{l}{\rule{0pt}{1em}\textsuperscript{\S} The average phenotype value for all genotypes in the test.}\\
\multicolumn{9}{l}{\rule{0pt}{1em}\textsuperscript{\P} The average value of the checks in the test.}\\
\end{tabular}}
\end{table}

@@

\begin{figure}[H]
\centering
\includegraphics[width=10 cm]{C:/Users/Jay/Documents/R/Projects/Jay_yield/exports/plots/example_plot.pdf}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}
